Lynott et al. (2014) - Potential error in statistics reporting

### 2017-11-12 manuscript.Rmd
Changed plot order in Figure 1, changed all instances of rounding to printnum() or printp(); minor formatting fixes

```{r echo=FALSE, fig.width=5, fig.height=7, fig.cap="Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests. **A)** Equivalence test: Tests the hypothesis ($H_0$) that an effect is smaller than $\\Delta_{L}$ *or* larger than $\\Delta_{U}$. **B)** NHST: Tests the hypothesis ($H_0$) that an effect is equal to 0. **C)** Minimal effects test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta_{L}$ *and* smaller than $\\Delta_{U}$. **D)** Inferiority test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta$."}

@@

@Norman2004 used a wide range of methods to determine the minimally important (or detectable) difference in measures for the quality of life in chronic diseases, and observe a surprisingly consistent minimally important difference of half a standard deviation, or d = 0.5. They make the argument that this bound of approximately half a standard deviation could be related to the limits of human discriminability. Therefore, even though a SESOI of d = 0.5 might seem quite large, it could be a plausible equivalence bound when examining effects that should be noticeable by single individuals.

@@

We distinguish between three categories of subjective justifications for the SESOI. First, researchers can use benchmarks. For example, one might set the SESOI to a standardized effect size of d = 0.5, which would allow one to reject effect as large or larger than a 'medium' effect size [@Cohen1988]. Similarly, effect sizes smaller than a Cohen's d of 0.1 are sometimes considered trivially small [@Maxwell2015]. Relying on a benchmark is the weakest possible justification of a SESOI, and should be avoided. Based on a review of 112 meta-analyses, @Weber2012 conclude that setting a SESOI in communication research to a medium effect size (*r* = 0.3, or d = 0.5) corresponds to rejecting the highest 25% of effect sizes reported in communications research, and @Hemphill2003 suggests that in psychology a SESOI of d = 0.5 would imply rejecting effects as large as the upper 33% of effect sizes reported in the psychological literature.

@@

Second, researchers can determine the SESOI based on previous work in the literature. Ideally, researchers would specify their SESOI in studies they report, but this is not yet common practice. It is thus up to researchers who build on earlier work to decide which effect size is too small to be meaningful, given an earlier study. @Simonsohn2015 recently proposed to set the SESOI to the effect size that an earlier study on the same effect would have had 33% power to detect. With this so-called 'small telescopes' approach, the equivalence bounds are thus determined primarily based on the sample size in the original study. For example, consider a study in which `r N` participants answered a question, and the results were analyzed with a one-sample t-test. For a two-sided test with an alpha of `r alpha`, this test had 33% power to detect an effect of d = `r round(pwr.t.test(n=N,sig.level=alpha,power=0.33,type="one.sample",alternative="two.sided")$d,3)`. Another example of how previous research can be used to determine the SESOI can be found in @Kordsmeyer2017, who defined the SESOI based on the mean of effect sizes reported in the literature. Thus, they examined whether their replication study could reject effects as large or larger than had on average been reported in the literature. Given random variation and bias in the literature, a more conservative approach could be to use a lower bound of a confidence interval around the meta-analytic effect size estimate [cf. @Perugini2014].

@@

Third, researchers can set the SESOI based on the sample size they are planning to collect. The amount of data you can collect limits the inferences you can make. Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect.^[This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a *p*-value *larger than* 0.05 in a traditional NHST. However, @Meyners2012 explains this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and the approach does not accurately control error rates.] This approach can be used when there are no quantitative predictions made by a theory, and no previous studies have been performed, but researchers *can* justify their sample size. In essence, this approach allows researchers to reject the presence of effects that can be reliably detected given a specific sample size. For example, a researcher who plans to perform a (two-sided) one-sample *t*-test using an alpha of 5%, based on data from 100 observations, has 90% power to detect an effect of d = 0.364. Using equivalence bounds of $\Delta_{L} = -0.364$ and $\Delta_{U} = 0.364$ in an equivalence test allows one to reject effects the study had high statistical power to detect. An equivalence test based on this approach does not answer any theoretical question (after all, the equivalence bounds are not based on any theoretical predictions) but it answers a *resource* question: Can effects be rejected that could reliably be detected given a specific sample size? In this case, concluding equivalence would suggest that effects which could reasonably be detected with a sample size of 100 can be rejected, and that future studies probably need larger sample sizes to examine a specific question. Whether or not the answer such an equivalence test provides is interesting depends on the sample sizes used in the study, and in part on how typical the chosen sample size is for the research area. An equivalence test based on 15 observations testing against a SESOI of d = 1 cannot be expected to substantially increase our knowledge, given that most effects in psychology are substantially smaller than these bounds and require larger sample sizes to be detected. By transparently reporting the effects one can detect and reject, based on the study design, researchers can communicate the information their study contributes, and provide a starting point for a discussion about what a reasonable SESOI is. 
```{r}
p <- 0.05
N <- 100
crit_d <- abs(qt(p/2, (N*2)-2))/sqrt(N/2)
```

@@

Equivalence bounds can be set in terms of standardized effect sizes, or raw differences. For example, the SESOI might be a raw mean difference of 0.5 points on a 7-point scale, or a Cohen’s d of 0.5. The main difference is that equivalence bounds set in raw differences are independent of the standard deviation, while equivalence bounds set as standardized effects are dependent on the standard deviation (since they are calculated as [FORMULA (X1-X2)/SD]. The observed standard deviation randomly varies across samples. In practice, this means that when you use standardized differences as bounds (e.g., d = 0.35) the equivalence test depends on the standard deviation in your the sample. The equivalence bound for a raw mean difference of 0.5 equals a standardized equivalence bound of d = 0.5 when the standard deviation is 1, but a standardized equivalence bound of d = 1 when the standard deviation is 0.5.

@@

@Banerjee2012 reported that `r orig.N` participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r orig.p`$, $d= `r orig.d`$). A close replication by @Brandt2014 with `r rep.n1+rep.n2` participants found no significant effect ($t(`r rep.df`)=`r rep.t`$, $p=`r rep.p`$, $d=`r rep.d`$). Following the small telescopes approach, we can calculate the effect size the original study had 33% power to detect --- $d = `r round(d.33, 2)`$ --- and use this as our SESOI. If we run a TOST with Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r round(-d.33, 2)`$ and ${\Delta}_U=`r round(d.33, 2)`$, we indeed find that the effect reported by the replication study is statistically equivalent, $t(`r round(Brandt$TOST_df, 2)`)=`r round(min(Brandt$TOST_t1, Brandt$TOST_t2), 2)`$, $p=`r round(max(Brandt$TOST_p1, Brandt$TOST_p2), 3)`$ (Figure 2A).

@@

@Hyde2008 report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, where the authors specify a trivial difference as an effect size smaller than d = 0.1. For example, in grade 3 the difference is d = 0.04, with a standard error of 0.002. When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of 0.005 to correct for multiple comparisons), using equivalence bounds of d = -0.1 and d = 0.1, we see that indeed, all effect size estimates are measured with such high precision, we can conclude that they fall within the equivalence bound of d = -0.1 and d = 0.1, and can be considered trivially small, according to the bounds set by the authors (figure 2B). However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions, and samples sizes are huge. This is one way in which the use of equivalence tests improve hypothesis testing procedures in science, by distinguishing between statistical significance and practical significance.

@@

@Moon2014, replicating @Shih1999, conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity. They observed a slightly reversed difference between the Asian primed group (*n* = 53, *M* = 0.46, *SD* = 0.17) and the control (*n* = 48, *M* = 0.50, *SD* - 0.18) which was not significant, d = -0.23, *t*(96.62) = 1.15, *p* = 0.26, two-sided). A non-significant null-hypothesis test leaves two possible conclusions:

@@

Let us imagine that we consult a teacher about the test, and discover that that grades for this test are set at every 6.25% increase in correct answers (F = 0% to 6.25% ... A+ = 93.75% to 100%). We can decide that we are only interested in test differences so far as they correspond to at least a 1 grade point increase or decrease. Thus, our SESOI becomes a difference in raw scores of 6.25%, or 0.0625.
We then calculate the TOST for a two-sample Welch’s t-test, assuming an alpha of 0.05 (two-sided), and using $\pm$ the SESOI of 0.0625 as our equivalence bounds ($\Delta$).
We find that the TOST is non-significant, t(`r round(Moon$TOST_df, 2)`) = `r round(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))), 2)`, p = `r round(max(Moon$TOST_p1, Moon$TOST_p2), 2)` (figure 2C). Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.^[In this example, the bounds of the TOST was set based on a raw effect of 6.25%. We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d ($\Delta = \pm `r round(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2), 2)`)$. Since this is a replication of a previous study, we could also have chosen to calculate the Cohen's d that corresponds to the raw SESOI in the *original* study, and used this as the bounds for the replication. Now we would no longer be answering a question about a raw effect independent of the samples. Instead, we are asking whether we in the replication can reject standardized effects larger than a certain standardized effect in the original study. Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d (and thus the bounds) would also be only slightly smaller ($\Delta = \pm `r round(0.0625 / sqrt((0.17^2 + 0.20^2) / 2), 2)`)$. As standard deviations of the original study compared to the replication become larger, the bounds become more narrow.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions. Further studies are needed, which can be analyzed using a (small-scale) meta-analysis. The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both. Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L}$ = -0.1 and $\Delta_{U}$ = 0.1).

@@

@Lynott2014 conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on subsequent judgments related to interpersonal warmth and prosocial behavior [replicating @Williams2008]. They observed that `r round(selfish.cold*100,2)`\% of participants who received a cold pack ($n_{cold}=`r round(n.cold,2)`$) opted to receive a reward for themselves, while `r round(selfish.hot*100,2)`\% of participants who received a warm pack ($n_{hot}=`r n.hot`$) did the same. In a z-test for the difference between the proportions, this effect is not statistically significant ($\mathit{Diff}=`r round(Lynott$dif * 100, 2)`$\%, $Z=`r round(Lynott$NHST_z,2)`$, $p=`r round(Lynott$NHST_p,2)`$). Having obtained a non-significant result, we are left with the same two  alternatives as in example 3.

In this case, it is not really clear what an objective criterion for the smallest effect size of interest could be. However, since this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could detect. This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = \pm `r round(qnorm(1-0.05/2),2)`$). To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error:

Having determined the SESOI, we may also decide that since this is a replication, we are only interested in whether the effect of receiving a cold pack is significantly *lower* than the smallest detectable effect of the original study. In this case, we can still use a TOST to analyse the data, but we only need to consider the test against $\Delta_{U}$ because an effect with CIs crossing $\Delta_{L}$ can still be *lower* than the original effect. 
We find that the test against $\Delta_{U}$ is `r ifelse(Lynott$TOST_p2 < Lynott$alpha, "significant", "non-significant")`, $z =$ `r round(Lynott$TOST_z2, 2)`, $p =$ `r ifelse(round(Lynott$TOST_p2, 3) >= 0.001, paste("=", round(Lynott$TOST_p2, 2)), "< 0.001")` (see figure 2D). Thus, we can conclude that the statistically non-significant effect is also statistically inferior to our SESOI.

@@

@Kahane2015 find no significant correlation between the perceived wrongness of a utilitarian choice in the classical sacrificial dilemmas and the new "greater good" dilemmas, $r(`r Kahane.N-2`) = `r round(Kahane.r.GG, 2)`$, $p = `r round(Kahane.p, 3)`$ ($N = `r Kahane.N`$^[Study 4 in @Kahane2015 had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of $N = `r Kahane.N`$.]). They conclude that the classical vignettes fail to capture "true"" utilitarianism, but also grant that this conclusion is dependent on the power of their study and that better-powered future studies might overturn their verdict: "Thus, while we cannot rule out the possibility that such an association could emerge in future studies using an even larger number of subjects or different measures, we submit that, in light of the present results, a robust association between ‘utilitarian’ judgment and genuine concern for the greater good seems extremely unlikely." (p. 206).
This inference --- that the result would be surprising if there was a true effect of a size the study could have detected --- can be formalized with an equivalence test for correlations and bounds set to an effect size the study had reasonable power to detect: With `r Kahane.N` participants, the study had 80% power to detect effects of $r = `r round(Kahane.r.80, 2)`$. Given bounds of $\Delta_{L} = `r -round(Kahane.r.80, 2)`$ and $\Delta_{U} = `r round(Kahane.r.80, 2)`$, we find that the result is indeed statistically equivalent, $r(`r Kahane.N-2`) = `r round(Kahane.r.GG, 2)`$, $p = `r round(max(Kahane$TOST_p1, Kahane$TOST_p2), 3)`$. This means that values smaller than $r = `r -round(Kahane.r.80, 2)`$ and larger than $r = `r round(Kahane.r.80, 2)`$ can be rejected at an alpha level of 5% and used as an initial benchmark which future studies could challenge using larger samples and more narrow equivalence bounds.

\begin{textbox}
\caption{Calculating an equivalence test in R}
\begin{framed}
{
\small
\setstretch{1}
If you want to calculate an equivalence test based on summary statistics, you can use the ``TOSTER'' package in the open-source programming language R. Using TOSTER requires very little experience with R, or programming in general. Below are the few simple steps required to calculate an equivalence test with TOSTER. 
\begin{enumerate}
\raggedright
  \setlength{\parskip}{1pt}
\item Download, install, and open R.
\item Run \verb!install.packages("TOSTER")! in the console.
\item Run \verb!library(TOSTER)! in the console.
\item The example below reproduces the result of example 1 in R. The parameters of the test are defined inside the parentheses, and you can run \verb!?TOSTtwo! in the console to get information about them. Simply copy the example to the console, replace the values with the corresponding values of your own study, and run the code. Results and a plot will automatically be printed.
\begin{verbatim}
TOSTtwo(m1  = 4.7857, 
        m2  = 4.6569,
        sd1 = 1.0897, 
        sd2 = 1.1895, 
        n1 = 49, 
        n2 = 51, 
        low_eqbound_d = -0.6401696, 
        high_eqbound_d = 0.6401696, 
        alpha = 0.05, 
        var.equal = FALSE)
\end{verbatim}
\item The example above uses the \verb!TOSTtwo()! function in the TOSTER package which calculates an equivalence test for an independent samples t-test. If you need to calculate TOST for a different type of test (e.g. a correlation), you can find documentation for \verb!TOSTr()! and other functions of the TOSTER package by going to {\sffamily \url{https://cran.r-project.org/web/packages/TOSTER/TOSTER.pdf}} in your web browser. 
\end{enumerate}}
\end{framed} 
\end{textbox}

@@

As with any statistical test, the usefulness of the result of an equivalence test depends on the question we ask. The question manifests itself in the bounds we set: Is our data surprising if the true effect lies outside of the bounds? If we test against very wide bounds, observing statistical significance can hardly be considered surprising, given that most effects in psychology are small to medium [@Hemphill2003]. Examining the papers citing @Lakens2017a, we see some researchers state a SESOI without providing a justification for it [@Brown2017; @Schumann2017].^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.] An equivalence test using a SESOI of d = 0.5 might very well answer a question the researchers are interested in [for one possible justification based on minimally important differences, see @Norman2004], but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.

>>>>

```{r echo=FALSE, fig.width=5, fig.height=7, fig.cap="Illustration of null hypotheses ($H_0$) and alternative hypotheses ($H_1$) for different types of significance tests. **A)** NHST: Tests the hypothesis ($H_0$) that an effect is equal to 0. **B)** Minimal effects test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta_{L}$ *and* smaller than $\\Delta_{U}$. **C)** Equivalence test: Tests the hypothesis ($H_0$) that an effect is smaller than $\\Delta_{L}$ *or* larger than $\\Delta_{U}$. **D)** Inferiority test: Tests the hypothesis ($H_0$) that an effect is larger than $\\Delta$."}

@@

@Norman2004 used a wide range of methods to determine the minimally important (or detectable) difference in measures for the quality of life in chronic diseases, and observe a surprisingly consistent minimally important difference of half a standard deviation, or $d = 0.5$. They make the argument that this bound of approximately half a standard deviation could be related to the limits of human discriminability. Therefore, even though a SESOI of $d = 0.5$ might seem quite large, it could be a plausible equivalence bound when examining effects that should be noticeable by single individuals.

@@

We distinguish between three categories of subjective justifications for the SESOI. First, researchers can use benchmarks. For example, one might set the SESOI to a standardized effect size of $d = 0.5$, which would allow one to reject effect as large or larger than a 'medium' effect size [@Cohen1988]. Similarly, effect sizes smaller than a Cohen's d of 0.1 are sometimes considered trivially small [@Maxwell2015]. Relying on a benchmark is the weakest possible justification of a SESOI, and should be avoided. Based on a review of 112 meta-analyses, @Weber2012 conclude that setting a SESOI in communication research to a medium effect size ($r = 0.3$, or $d = 0.5$) corresponds to rejecting the highest 25% of effect sizes reported in communications research, and @Hemphill2003 suggests that in psychology a SESOI of $d = 0.5$ would imply rejecting effects as large as the upper 33% of effect sizes reported in the psychological literature.

@@

Second, researchers can determine the SESOI based on previous work in the literature. Ideally, researchers would specify their SESOI in studies they report, but this is not yet common practice. It is thus up to researchers who build on earlier work to decide which effect size is too small to be meaningful, given an earlier study. @Simonsohn2015 recently proposed to set the SESOI to the effect size that an earlier study on the same effect would have had 33% power to detect. With this so-called 'small telescopes' approach, the equivalence bounds are thus determined primarily based on the sample size in the original study. For example, consider a study in which `r N` participants answered a question, and the results were analyzed with a one-sample t-test. For a two-sided test with an alpha of `r alpha`, this test had 33% power to detect an effect of $d = `r printnum(pwr.t.test(n=N,sig.level=alpha,power=0.33,type="one.sample",alternative="two.sided")$d)`$. Another example of how previous research can be used to determine the SESOI can be found in @Kordsmeyer2017, who defined the SESOI based on the mean of effect sizes reported in the literature. Thus, they examined whether their replication study could reject effects as large or larger than had on average been reported in the literature. Given random variation and bias in the literature, a more conservative approach could be to use a lower bound of a confidence interval around the meta-analytic effect size estimate [cf. @Perugini2014].

@@

Third, researchers can set the SESOI based on the sample size they are planning to collect. The amount of data you can collect limits the inferences you can make. Given the alpha level and the sample size, researchers can calculate the smallest effect size that they have sufficient power to detect.^[This approach is conceptually very similar to the *power approach*, where the effect size you had 95% power to detect is calculated, and the presence of effects more extreme than this value is rejected after observing a *p*-value *larger than* 0.05 in a traditional NHST. However, @Meyners2012 explains this approach is not recommended (even though it is common) because it ignores the possibility that effects are significant *and* equivalent, and the approach does not accurately control error rates.] This approach can be used when there are no quantitative predictions made by a theory, and no previous studies have been performed, but researchers *can* justify their sample size. In essence, this approach allows researchers to reject the presence of effects that can be reliably detected given a specific sample size. For example, a researcher who plans to perform a (two-sided) one-sample *t*-test using an alpha of 5%, based on data from 100 observations, has 90% power to detect an effect of $d = 0.364$. Using equivalence bounds of $\Delta_{L} = -0.364$ and $\Delta_{U} = 0.364$ in an equivalence test allows one to reject effects the study had high statistical power to detect. An equivalence test based on this approach does not answer any theoretical question (after all, the equivalence bounds are not based on any theoretical predictions) but it answers a *resource* question: Can effects be rejected that could reliably be detected given a specific sample size? In this case, concluding equivalence would suggest that effects which could reasonably be detected with a sample size of 100 can be rejected, and that future studies probably need larger sample sizes to examine a specific question. Whether or not the answer such an equivalence test provides is interesting depends on the sample sizes used in the study, and in part on how typical the chosen sample size is for the research area. An equivalence test based on 15 observations testing against a SESOI of $d = 1$ cannot be expected to substantially increase our knowledge, given that most effects in psychology are substantially smaller than these bounds and require larger sample sizes to be detected. By transparently reporting the effects one can detect and reject, based on the study design, researchers can communicate the information their study contributes, and provide a starting point for a discussion about what a reasonable SESOI is. 
```{r}
p <- 0.05
N <- 100
crit_d <- abs(qt(p/2, (N*2)-2))/sqrt(N/2)
```

@@

Equivalence bounds can be set in terms of standardized effect sizes, or raw differences. For example, the SESOI might be a raw mean difference of 0.5 points on a 7-point scale, or a Cohen’s d of 0.5. The main difference is that equivalence bounds set in raw differences are independent of the standard deviation, while equivalence bounds set as standardized effects are dependent on the standard deviation (since they are calculated as [FORMULA (X1-X2)/SD]. The observed standard deviation randomly varies across samples. In practice, this means that when you use standardized differences as bounds (e.g., $d = 0.35$) the equivalence test depends on the standard deviation in your the sample. The equivalence bound for a raw mean difference of 0.5 equals a standardized equivalence bound of $d = 0.5$ when the standard deviation is 1, but a standardized equivalence bound of $d = 1$ when the standard deviation is 0.5.

@@

@Banerjee2012 reported that `r orig.N` participants who had been asked to describe an unethical deed from their past judged the room to be darker than participants who had been asked to describe an ethical deed ($M_{unethical}= `r orig.m1`$, ${SD}_{unethical}= `r orig.sd1`$, $M_{ethical}=`r orig.m2`$, ${SD}_{ethical}=`r orig.sd2`$, $t(`r orig.df`)= `r orig.t`$, $p= `r printp(orig.p)`$, $d= `r printnum(orig.d)`$). A close replication by @Brandt2014 with `r rep.n1+rep.n2` participants found no significant effect ($t(`r printnum(rep.df)`)=`r printnum(rep.t)`$, $p=`r printp(rep.p)`$, $d=`r printnum(rep.d)`$). Following the small telescopes approach, we can calculate the effect size the original study had 33% power to detect --- $d = `r printnum(d.33)`$ --- and use this as our SESOI. If we run a TOST with Welch's *t*-test for independent samples and equivalence bounds of ${\Delta}_L=`r printnum(-d.33)`$ and ${\Delta}_U=`r printnum(d.33)`$, we indeed find that the effect reported by the replication study is statistically equivalent, $t(`r printnum(Brandt$TOST_df)`)=`r printnum(min(Brandt$TOST_t1, Brandt$TOST_t2))`$, $p=`r printp(max(Brandt$TOST_p1, Brandt$TOST_p2))`$ (Figure 2A).

@@

@Hyde2008 report that effect sizes for gender differences in mathematics tests across the 7 million students in the US represent trivial differences, where the authors specify a trivial difference as an effect size smaller than $d = 0.1$. For example, in grade 3 the difference is $d = 0.04$, with a standard error of 0.002. When we perform equivalence tests on the meta-analytic effect sizes of IQ difference for grades 2 to 11 (using an alpha level of 0.005 to correct for multiple comparisons), using equivalence bounds of $d = -0.1$ and $d = 0.1$, we see that indeed, all effect size estimates are measured with such high precision, we can conclude that they fall within the equivalence bound of $d = -0.1$ and $d = 0.1$, and can be considered trivially small, according to the bounds set by the authors (figure 2B). However, note that all of the effects are also statistically different from zero, as one might expect when there is no random assignment to conditions, and samples sizes are huge. This is one way in which the use of equivalence tests improve hypothesis testing procedures in science, by distinguishing between statistical significance and practical significance.

@@

@Moon2014, replicating @Shih1999, conducted a study to investigate whether Asian-American women would perform better on a maths test when primed with their Asian identity. They observed a slightly reversed difference between the Asian primed group ($n = 53$, $M = 0.46$, $SD = 0.17$) and the control ($n = 48$, $M = 0.50$, $SD = 0.18$) which was not significant, $d = -0.23$, $t(96.62) = 1.15$, $p = 0.26$ (two-sided). A non-significant null-hypothesis test leaves two possible conclusions:

@@

Let us imagine that we consult a teacher about the test, and discover that that grades for this test are set at every 6.25% increase in correct answers (F = 0% to 6.25% ... A+ = 93.75% to 100%). We can decide that we are only interested in test differences so far as they correspond to at least a 1 grade point increase or decrease. Thus, our SESOI becomes a difference in raw scores of 6.25%, or 0.0625.
We then calculate the TOST for a two-sample Welch’s t-test, assuming an alpha of 0.05 (two-sided), and using $\pm$ the SESOI of 0.0625 as our equivalence bounds ($\Delta$).
We find that the TOST is non-significant, t(`r printnum(Moon$TOST_df)`) = `r printnum(min(abs(c(Moon$TOST_t1, Moon$TOST_t2))))`, p = `r printp(max(Moon$TOST_p1, Moon$TOST_p2))` (figure 2C). Thus, we cannot reject the possibility that the true effect really is larger than our SESOI.^[In this example, the bounds of the TOST was set based on a raw effect of 6.25%. We could have converted this raw effect into a standardized effect (Cohen's d) for the current sample, and we would have gotten the exact same result, only then in Cohen's d ($\Delta = \pm `r printnum(0.0625 / sqrt((sd.asian^2 + sd.control^2) / 2))`)$. Since this is a replication of a previous study, we could also have chosen to calculate the Cohen's d that corresponds to the raw SESOI in the *original* study, and used this as the bounds for the replication. Now we would no longer be answering a question about a raw effect independent of the samples. Instead, we are asking whether we in the replication can reject standardized effects larger than a certain standardized effect in the original study. Since the standard deviations of the original study were only slightly higher than in the replication in this case, Cohen's d (and thus the bounds) would also be only slightly smaller ($\Delta = \pm `r printnum(0.0625 / sqrt((0.17^2 + 0.20^2) / 2))`)$. As standard deviations of the original study compared to the replication become larger, the bounds become more narrow.] When results are neither statistically different from zero nor statistically equivalent, there is insufficient data to draw conclusions. Further studies are needed, which can be analyzed using a (small-scale) meta-analysis. The additional data will narrow the confidence interval around the observed effect, allowing us to reject the null, reject the SESOI, or both. Analogous to the large sample sizes needed to detect small effects, a lot of data is needed to reject the SESOI when very narrow bounds are used for the equivalence test (e.g., $\Delta_{L} = -0.1$ and $\Delta_{U} = 0.1$).

@@

@Lynott2014 conducted a study to investigate the effect of being exposed to physically warm or cold stimuli on subsequent judgments related to interpersonal warmth and prosocial behavior [replicating @Williams2008]. They observed that `r printnum(selfish.cold*100)`\% of participants who received a cold pack ($n_{cold}=`r n.cold`$) opted to receive a reward for themselves, while `r printnum(selfish.hot*100)`\% of participants who received a warm pack ($n_{hot}=`r n.hot`$) did the same. In a z-test for the difference between the proportions, this effect is not statistically significant ($\mathit{Diff}=`r printnum(Lynott$dif * 100)`$\%, $Z=`r printnum(Lynott$NHST_z)`$, $p=`r printp(Lynott$NHST_p)`$). Having obtained a non-significant result, we are left with the same two  alternatives as in example 3.

In this case, it is not really clear what an objective criterion for the smallest effect size of interest could be. However, since this is a replication, we could decide that we do not care about an effect if the difference is smaller than the smallest effect that the original study could detect. This essentially means that we use the critical z value (~1.96 in a two-tailed test with an alpha of 0.05) as our bounds ($\Delta = \pm `r printnum(qnorm(1-0.05/2))`$). To figure out what difference corresponds to a critical z in the original study, we multiply critical z with the standard error:

Having determined the SESOI, we may also decide that since this is a replication, we are only interested in whether the effect of receiving a cold pack is significantly *lower* than the smallest detectable effect of the original study. In this case, we can still use a TOST to analyse the data, but we only need to consider the test against $\Delta_{U}$ because an effect with CIs crossing $\Delta_{L}$ can still be *lower* than the original effect. 
We find that the test against $\Delta_{U}$ is `r ifelse(Lynott$TOST_p2 < Lynott$alpha, "significant", "non-significant")`, $z =$ `r printnum(Lynott$TOST_z2)`, $p `r printp(Lynott$TOST_p2)`$ (see figure 2D). Thus, we can conclude that the statistically non-significant effect is also statistically inferior to our SESOI.

@@

@Kahane2015 find no significant correlation between the perceived wrongness of a utilitarian choice in the classical sacrificial dilemmas and the new "greater good" dilemmas, $r(`r Kahane.N-2`) = `r printnum(Kahane.r.GG)`$, $p = `r printp(Kahane.p)`$ ($N = `r Kahane.N`$^[Study 4 in @Kahane2015 had a final sample size of $N = `r nrow(Kahane.data)`$, but due to missing data in `r sum(is.na(Kahane.data))` case, the correlation reported here is based on a sample of $N = `r Kahane.N`$.]). They conclude that the classical vignettes fail to capture "true" utilitarianism, but also grant that this conclusion is dependent on the power of their study and that better-powered future studies might overturn their verdict: "Thus, while we cannot rule out the possibility that such an association could emerge in future studies using an even larger number of subjects or different measures, we submit that, in light of the present results, a robust association between ‘utilitarian’ judgment and genuine concern for the greater good seems extremely unlikely." (p. 206).
This inference --- that the result would be surprising if there was a true effect of a size the study could have detected --- can be formalized with an equivalence test for correlations and bounds set to an effect size the study had reasonable power to detect: With `r Kahane.N` participants, the study had 80% power to detect effects of $r = `r printnum(Kahane.r.80)`$. Given bounds of $\Delta_{L} = `r printnum(-Kahane.r.80)`$ and $\Delta_{U} = `r printnum(Kahane.r.80)`$, we find that the result is indeed statistically equivalent, $r(`r Kahane.N-2`) = `r printnum(Kahane.r.GG)`$, $p = `r printp(max(Kahane$TOST_p1, Kahane$TOST_p2))`$. This means that values smaller than $r = `r printnum(-Kahane.r.80)`$ and larger than $r = `r printnum(Kahane.r.80)`$ can be rejected at an alpha level of 5% and used as an initial benchmark which future studies could challenge using larger samples and more narrow equivalence bounds.

\begin{textbox}
\caption{Calculating an equivalence test in R.}
\begin{framed}
{
\small
\setstretch{1}
If you want to calculate an equivalence test based on summary statistics, you can use the ``TOSTER'' package in the open-source programming language R. Using TOSTER requires very little experience with R, or programming in general. Below are the few simple steps required to calculate an equivalence test with TOSTER. 
\begin{enumerate}
\raggedright
  \setlength{\parskip}{1pt}
\item Download, install, and open R.
\item Run \verb!install.packages("TOSTER")! in the console.
\item Run \verb!library(TOSTER)! in the console.
\item The example below reproduces the result of example 1 in R. The parameters of the test are defined inside the parentheses, and you can run \verb!?TOSTtwo! in the console to get information about them. Simply copy the example to the console, replace the values with the corresponding values of your own study, and run the code. Results and a plot will automatically be printed.
\begin{verbatim}
TOSTtwo(m1  = 4.7857, 
        m2  = 4.6569,
        sd1 = 1.0897, 
        sd2 = 1.1895, 
        n1 = 49, 
        n2 = 51, 
        low_eqbound_d = -0.6401696, 
        high_eqbound_d = 0.6401696, 
        alpha = 0.05, 
        var.equal = FALSE)
\end{verbatim}
\item The example above uses the \verb!TOSTtwo()! function in the TOSTER package which calculates an equivalence test for an independent samples t-test. If you need to calculate TOST for a different type of test (e.g. a correlation), you can find documentation for \verb!TOSTr()! and other functions of the TOSTER package by going to {\sffamily \url{https://cran.r-project.org/web/packages/TOSTER/TOSTER.pdf}} in your web browser. 
\end{enumerate}}
\end{framed} 
\end{textbox}

@@

As with any statistical test, the usefulness of the result of an equivalence test depends on the question we ask. The question manifests itself in the bounds we set: Is our data surprising if the true effect lies outside of the bounds? If we test against very wide bounds, observing statistical significance can hardly be considered surprising, given that most effects in psychology are small to medium [@Hemphill2003]. Examining the papers citing @Lakens2017a, we see some researchers state a SESOI without providing a justification for it [@Brown2017; @Schumann2017].^[In these two cases, one could perhaps guess that benchmark values for a Cohen’s d of “medium” size was chosen.] An equivalence test using a SESOI of $d = 0.5$ might very well answer a question the researchers are interested in [for one possible justification based on minimally important differences, see @Norman2004], but researchers should always provide a detailed justification of their chosen SESOI, given that this is in essence a justification for why they ask a specific question from their data.
